figures?

### 2024-01-05 Updates/2023-11-17 figures_draft.Rmd
recaulcute skill and save plots

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = crps_climatology - crps,
         skill_logs = logs_climatology - logs,
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1,
         model_id != 'fARIMA') |> 
  arrange(reference_datetime, site_id, datetime, model_id) 
```

@@

Top 10 models and all models coloured by type and showing if they include covariates or not. 
```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  slice_max(mean_crps, n= 10, na_rm = T) |>
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.04,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw() + 
    guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                  title = 'Uses weather covariates?'),
           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

all_temp <- temperature_skill |> 
    group_by(model_id) |> 
    summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
    left_join(model_meta) |> 
    mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
    ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
    geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                     stat = "identity",
                     alpha = 0.8, 
                     pattern_fill = "black",
                     colour = "black", 
                     pattern_spacing = 0.03,
                     pattern_alpha = 1) + 
    scale_y_reordered() +
    labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
    geom_vline(xintercept = 0) +
    scale_fill_manual(values = cols_modeltype) +
    scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
    scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
    theme_bw() + 
    guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, 
                  common.legend = T,
                  legend = 'top', labels = 'auto')

@@

```{r model_type-site}
temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
  labs(y='model_id', x = 'mean CRPS score') +
  scale_colour_manual(values = cols_modeltype) +
  theme_bw()

temperature_climatology |> 
  group_by(site_id) |> 
  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
  arrange(mean_score)
```

@@

A <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'H') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = model_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw()

ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2)
```

@@

sd <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_sd = mean(sd, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_sd, colour = model_id)) +
  geom_line()  +  
  labs(y='mean standard deviation', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw()

bias <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
         abs_bias = abs(mean - observation)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_abs_bias, colour = model_id)) +
  geom_line()  +  
  labs(y='mean absolute bias (oC)', x = 'horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 2.5))

ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2)
```

>>>>

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = 1 - (crps/crps_climatology),
         skill_logs = 1 - (logs/logs_climatology),
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1,
         model_id != 'fARIMA') |> 
  arrange(reference_datetime, site_id, datetime, model_id) 
```

@@

Top 10 models and all models coloured by type and showing if they include covariates or not. 
```{r model-type, fig.width=10}
top10_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |>
  slice_max(mean_crps, n= 10, na_rm = T) |>
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.04,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe')) +
  scale_pattern_type_manual(values=c(NA, NA)) +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                title = 'Uses weather covariates?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

all_temp <- 
  temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps),
         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
  # geom_rect(aes(ymin = 'flareGOTM_noDA', 
  #               ymax ='flareGLM', 
  #               xmin = -Inf, 
  #               xmax = Inf)) +
  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
                   stat = "identity",
                   alpha = 0.8, 
                   pattern_fill = "black",
                   colour = "black", 
                   pattern_spacing = 0.03,
                   pattern_alpha = 1) + 
  scale_y_reordered() +
  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
  theme_bw() + 
  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))

ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, align = 'hv',
                  common.legend = T,
                  legend = 'top', labels = 'auto') |> 
  ggsave(filename = './Plots/Figure1.png', width = 25, height = 10, units = 'cm')

@@

```{r model_type-site}
clim_mean <- temperature_climatology |> 
  group_by(site_id) |> 
  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
  arrange(mean_score)

n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = -3.5) |> 
  filter(model_type != 'null')

(temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  # geom_point(data = clim_mean, aes(x = mean_score, y = site_id), colour = 'black') +
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
  labs(y='Site', x = 'mean CRPS skill (vs climatology)') +
  scale_colour_manual(values = cols_modeltype) +
  theme_bw()) |> 
  ggsave(filename = './Plots/Figure2.png', width = 15, height = 15, units = 'cm')

@@

A <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'H') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = model_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9, name = 'Model') +
  theme_bw()

ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2) |> 
  ggsave(filename = 'Plots/Figure3.png', width = 15, height = 17, units = 'cm')
```

@@

sd <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_sd = mean(sd, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_sd, colour = model_id)) +
  geom_line()  +  
  labs(y='mean standard deviation', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw()

bias <- temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
         abs_bias = abs(mean - observation)) |> 
  filter(model_id %in% top_mods$model_id, 
         horizon > 0, 
         horizon <=30) |>
  group_by(model_id, horizon) |> 
  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_abs_bias, colour = model_id)) +
  geom_line()  +  
  labs(y='mean absolute bias', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 2.5))

ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2) |> 
  ggsave(filename = 'Plots/Figure4.png')
```
